```{r, echo=FALSE, warning=FALSE, message=FALSE}
opts_chunk$set(echo=FALSE)
setwd("E:\\Reproducible Research\\Assessment 1")
options(digits=7)

library(plyr)
library(lubridate)
library(lattice)
library(xtable)

allData <- read.csv("./activity.csv", stringsAsFactors=FALSE)
activityData <- allData[complete.cases(allData),]
```

## What is mean total number of steps taken per day?

```{r TotalStepsNoNA}
plotData <-tapply(activityData$steps, activityData$date, FUN = sum)
barplot( plotData,
         col="mistyrose",
         main='Total number of steps each day (missing values removed)',
         xlab="Date",
         ylab="Steps",
         cex.names=0.5,
         names.arg=format(ymd(names(plotData)),"%d/%m")
      )
```

The mean total is `r sprintf("%.1f", mean(plotData))`.  The median total is `r median(plotData)`.

---

## What is the average daily activity pattern?

```{r DailyActivity}
plotData <- aggregate(steps~interval, data=activityData, FUN=mean)
plot( plotData$interval,
      plotData$steps, 
      main="Mean number of steps per 5-minute interval",
      xlab="Interval",
      ylab="Steps",
      type="l",
      col="darkcyan"
    )

intervalMeans <- aggregate(steps~interval, data=activityData, FUN=mean)
x<-intervalMeans[which.max(intervalMeans$steps),]$interval
```

Interval `r sprintf("%d", x)` on average across all the days in the dataset, contains the maximum number of steps.

---

## Inputing missing values

The total number of rows with 'NA's is `r nrow(allData) - nrow(activityData)`.

```{r TotalStepsWithNA}
allData$steps <- ifelse( is.na(allData$steps), 
                         intervalMeans$steps[ match(allData$interval, intervalMeans$interval) ], 
                         allData$steps )
plotData <-tapply(allData$steps, allData$date, FUN = sum)
barplot( plotData,
         col="mistyrose",
         main='Total number of steps each day (missing values estimated)',
         xlab="Date",
         ylab="Steps",
         cex.names=0.5,
         names.arg=format(ymd(names(plotData)),"%d/%m")
      )
```

The mean total is `r sprintf("%.1f", mean(plotData))`.  The median total is `r sprintf("%.1f", median(plotData))`.

As the NA replacement algorithm substituted the interval mean, the mean total has not changed but the median total has.

---

## Are there differences in activity patterns between weekdays and weekends?

```{r WeeklyActivity}
allData$date <- ymd(allData$date)
allData$weekend <- sapply( allData$date, 
                           function(x) 
                           { 
                              ifelse( wday(x, label=TRUE) %in% c('Sat', 'Sun'),
                                      "weekend", 
                                      "weekday" )
                           } 
                         )
plotData <- aggregate(steps~interval+weekend, data=allData, FUN=mean)

xyplot( steps~interval | weekend, 
        main='Mean number of steps per 5-minute interval',
        xlab="Interval",
        ylab="Steps",        
        type="l",
        col="darkcyan",
        layout=c(1,2),
        plotData )
```




